---
title: "GO_MWU_MN_pipeline"
author: "Melissa Naugle"
date: "11/20/2020"
output: html_document
---


```{r setup}
setwd("/Users/Melissa/Desktop/GitHub/RNAseq_allsites_Barshisreference/GO/GO_MWU/")
rm( list = ls())
graphics.off()
library(ggplot2)
library(ggpubr)
library(plyr)
library(dplyr)
library(reshape2)
library(ape)
library(data.table)
```

##To get started
#From https://github.com/z0on/GO_MWU/blob/master/README_GO_MWU.txt

1. Put all this into the same directory:
	- scripts: GO_MWU.R, gomwu_a.pl, gomwu_b.pl, gomwu.functions.R
	#downloaded from: https://github.com/z0on/GO_MWU
	
	- GO hierarchy file 
		(version 1.0, http://www.geneontology.org/GO.downloads.ontology.shtml)
		#downloaded from above website (go.obo)
		
	- table of GO annotations for your sequences: two-column (gene id - GO terms), 
		tab-delimited, one line per gene, multiple GO terms separated by semicolon. 
		If you have multiple lines per gene, use nrify_GOtable.pl to merge them.
		#created in this script
		
	- table of measure of interest for your sequences: two columns of comma-separated 
		values: gene id, continuous measure of significance (such as fold-change or 
		-log(p-value)). To perform standard GO enrichment analysis based on Fisher's 
		exact test, use binary measure (1 or 0, i.e., either significant or not).
		#use ref transcriptome and DE file 
		#used ALL DGES, maybe just try those between certain comparisons?
		#made in r script titled 'create_contig_signif_file_GOMWU.R' and gives file called contigs_signif_file_GOMWU.txt
		

2. Make sure you have perl and R. The R part requires package "ape", which 
you might need to install prior to running this method.

3. Open GO_MWU.R script; edit the input file names, mark and execute bits of code
separated by blank lines one by one. Follow instructions given as comments in the script.

4. Drag corner of the plot to rescale and match text and tree. After this, to
achieve better "word map" effect, rerun gomwuPlot with modified "txtsize" parameter.

5. Save the plot as pdf file.


#Create table of GO annotations

table of GO annotations for your sequences: two-column (gene id - GO terms), 
		tab-delimited, one line per gene, multiple GO terms separated by semicolon. 
		If you have multiple lines per gene, use nrify_GOtable.pl to merge them.
		
Using Barshis annotation file
Not sure where / how this was generated?

```{r g}
annot <- read.csv("33496_MasterCombinedAnnotationTable.csv", header = T)
head(annot)
#sep by " // " in GO column

GOtable <- annot %>%
  select(ContigName, GO)

GOtable$GO <- gsub(" // ", ";", GOtable$GO)
head(GOtable)

GOtable$GO <- gsub("No_GOcodes", "unknown", GOtable$GO)
GOtable$GO <- gsub("No_GO", "unknown", GOtable$GO)
head(GOtable)

write_delim(GOtable, path = "GO_annot_table", delim = "\t", col_names = F)
```

#To make signif file with log fc

```{r}

#read in file contigs of interest with common core genes 
contigs_heat_control <- read.csv("/Users/Melissa/Desktop/GitHub/RNAseq_allsites_Barshisreference/data/DE_comparison_lists/DE_middle_of_venn_classic_FDR0.05.csv")

#read in DE file 
DGE <- read.csv("../../GLM_edgeR_outputs/data/tagsTblANOVA.one_all_heat_v_control.csv", header = T)
DGE <- as.data.frame(DGE)
DGE <- DGE %>%
  select(1,2,4)
head(DGE)
nrow(DGE)

DGE$Ourcontigname <- row.names(DGE)
row.names(DGE) <- NULL

#select contigs of interest from DGE list
DGE_of_interest <- DGE[DGE$Ourcontigname%in%contigs_heat_control$ContigID,]


#read in file with all contigs from transciptome
allcontigs <- read.csv("33496_MasterCombinedAnnotationTable.csv", header = T)
nrow(allcontigs)
allcontigs <- allcontigs %>%
  select(1)
head(allcontigs)


contigs_and_signif <- merge(DGE_of_interest, allcontigs, all = T)
head(contigs_and_signif)
nrow(contigs_and_signif)

contigs_and_signif[is.na(contigs_and_signif)] <- 0

contigs_and_signif_fc <- contigs_and_signif %>%
  select(1,2)

contigs_and_signif_p <- contigs_and_signif %>%
  select(1,4)

head(contigs_and_signif_fc)
head(contigs_and_signif_p)

write.csv(contigs_and_signif_fc, "contigs_signif_file_GOMWU_all_heat_v_control_logfoldchange.csv", row.names = F, quote = F)

write.csv(contigs_and_signif_p, "contigs_signif_file_GOMWU_all_heat_v_control_p.csv", row.names = F, quote = F)

```





