---
title: "Venndiagram"
author: "Melissa Naugle"
date: "3/25/2021"
output: html_document
---

##For MSCI430 Class

#1. Set WD, load packages

```{r setup}
setwd(dir = "~/Desktop/GitHub/MSCI430_corals_Sp21/")
rm( list = ls())
graphics.off()
library(tidyverse) 
library(reshape2) 
library(ggthemes) 
library(VennDiagram)
library(dplyr)
```

#2. Read in data files

Here I have files with DEGs between control and treatment at each of the 5 sites
Could change these to to just look at upregulated or downregulated, or to a diff p value

```{r readdata}
#Read in data files, selecting only the contig name (ContigID) column

CoPtE_CoPtH <-read.csv("../data/EdgeR_P0.05/RSEM.isoform.counts.matrix.CoPt_E_vs_CoPt_H.edgeR.DE_results.P0.05_C2.DE.subset", sep = "\t")
CoPtE_CoPtH <- setNames(cbind(rownames(CoPtE_CoPtH), CoPtE_CoPtH, row.names = NULL), c("ContigID", colnames(CoPtE_CoPtH)))
head(CoPtE_CoPtH)
CoPtE_CoPtH <- select(CoPtE_CoPtH, "ContigID")

TeleE_TeleH <-read.csv("../data/EdgeR_P0.05/RSEM.isoform.counts.matrix.Ftele_E_vs_Ftele_H.edgeR.DE_results.P0.05_C2.DE.subset", sep = "\t")
TeleE_TeleH <- setNames(cbind(rownames(TeleE_TeleH), TeleE_TeleH, row.names = NULL), c("ContigID", colnames(TeleE_TeleH)))
head(TeleE_TeleH)
TeleE_TeleH <- select(TeleE_TeleH, "ContigID")

AluE_AluH <-read.csv("../data/EdgeR_P0.05/RSEM.isoform.counts.matrix.Falu_E_vs_Falu_H.edgeR.DE_results.P0.05_C2.DE.subset", sep = "\t")
AluE_AluH <- setNames(cbind(rownames(AluE_AluH), AluE_AluH, row.names = NULL), c("ContigID", colnames(AluE_AluH)))
head(AluE_AluH)
AluE_AluH <- select(AluE_AluH, "ContigID")

CanE_CanH <-read.csv("../data/EdgeR_P0.05/RSEM.isoform.counts.matrix.Can_E_vs_Can_H.edgeR.DE_results.P0.05_C2.DE.subset", sep = "\t")
CanE_CanH <- setNames(cbind(rownames(CanE_CanH), CanE_CanH, row.names = NULL), c("ContigID", colnames(CanE_CanH)))
head(CanE_CanH)
CanE_CanH <- select(CanE_CanH, "ContigID")

VatE_VatH <-read.csv("../data/EdgeR_P0.05/RSEM.isoform.counts.matrix.Vat_E_vs_Vat_H.edgeR.DE_results.P0.05_C2.DE.subset", sep = "\t")
VatE_VatH <- setNames(cbind(rownames(VatE_VatH), VatE_VatH, row.names = NULL), c("ContigID", colnames(VatE_VatH)))
head(VatE_VatH)
VatE_VatH <- select(VatE_VatH, "ContigID")

```

#3. Make categories for venn diagram

```{r make_venn_categories}
#Use intersect function to find contig names shared between comparisons. 

area1 <- CanE_CanH
area2 <- CoPtE_CoPtH
area3 <- AluE_AluH
area4 <- VatE_VatH
area5 <- TeleE_TeleH

n12 <- intersect(area1,area2, by="ContigID")
n13 <- intersect(area1,area3, by="ContigID")
n14 <- intersect(area1,area4, by="ContigID")
n15 <- intersect(area1,area5, by="ContigID")
n23 <- intersect(area2,area3, by="ContigID")
n24 <- intersect(area2,area4, by="ContigID")
n25 <- intersect(area2,area5, by="ContigID")
n34 <- intersect(area3,area4, by="ContigID")
n35 <- intersect(area3,area5, by="ContigID")
n45 <- intersect(area4,area5, by="ContigID")

n123 <- intersect(n12,area3, by="ContigID")
n124 <- intersect(n12,area4, by="ContigID")
n125 <- intersect(n12,area5, by="ContigID")
n134 <- intersect(n13,area4, by="ContigID")
n135 <- intersect(n13,area5, by="ContigID")
n145 <- intersect(n14,area5, by="ContigID")
n234 <- intersect(n23,area4, by="ContigID")
n235 <- intersect(n23,area5, by="ContigID")
n245 <- intersect(n24,area5, by="ContigID")
n345 <- intersect(n34,area5, by="ContigID")

n1234 <- intersect(n123,area4, by="ContigID")
n1235 <- intersect(n123,area5, by="ContigID")
n1245 <- intersect(n124,area5, by="ContigID")
n1345 <- intersect(n134,area5, by="ContigID")
n2345 <- intersect(n234,area5, by="ContigID")
n12345 <- intersect(n1234,area5, by-"ContigID")

write.csv(n12345, "../data/DE_middle_of_venn_classic_FDR0.05.csv", row.names = F)

```

#4. Make Venn Diagrams

This script makes 2 diagrams
1. 3 circle venn with Coconut Point, Faga'alu, and Faga'tele bw heat and control
2. 5 circle venn with all sites bw heat and control

Note: the venn.plot + grid.draw code doesn't work if you run it line by line, you need to run the entire chunk or highlight all the code and run at the same time 

Other options:
draw.single.venn() = normal venn diagram with 2 circles
draw.quad.venn() = 4 circle venn diagram

```{r make_venn}
#Plot Venn diagram with numbers generated above

#Can do three sites
venn.plot <- draw.triple.venn( 
  area1 = nrow(area1), area2 = nrow(area2), area3 = nrow(area3), 
  n12 = nrow(n12), n13 = nrow(n13), n23 = nrow(n23), n123 = nrow(n123), 
  #make sure this is in the correct order
  category = c("Cannery", "Coconut Point", "Faga'alu"), 
  #choose colors
  fill = c("skyblue", "mediumorchid", "salmon"), 
  #change line type (this means no line)
  lty = "blank", 
  #change size of category labels 
  cat.cex = 1.2,
  #change size of labels
  cex = 2
);

#Save the  Venn diagram
grid.newpage()
png("../Figures/venn_controlvheat_3sites_P0.05_04_26_2021.png")
grid.draw(venn.plot)
dev.off()

#or five sites!
venn.plot <- draw.quintuple.venn(
  area1 = nrow(area1), area2 = nrow(area2), area3 = nrow(area3), area4 = nrow(area4), area5 = nrow(area5),
  n12 = nrow(n12), n13 = nrow(n13), n14 = nrow(n14) , n15 = nrow(n15),
  n23 = nrow(n23), n24 = nrow(n24), n25 = nrow(n25), 
  n34 = nrow(n34), n35 = nrow(n35), n45 = nrow(n45), 
  n123 = nrow(n123), n124 = nrow(n124), n125 = nrow(n125), 
  n134 = nrow(n134), n135 = nrow(n135), n145 = nrow(n145), 
  n234 = nrow(n234), n235 = nrow(n235), n245 = nrow(n245), 
  n345 = nrow(n345), n1234 = nrow(n1234), n1235 = nrow(n1235),
  n1245 = nrow(n1245), n1345 = nrow(n1345), n2345 = nrow(n2345), n12345 = nrow(n12345), 
  category = c("Cannery", "Coconut Point", "Faga'alu", "Vatia", "Faga'tele"), 
  fill = c("salmon", "pink", "mediumorchid", "palegreen", "skyblue"), 
  #change line type (this is a solid line)
  lty = 1, 
  #change size of category labels
  cat.cex = 1.2,
  #this makes labels prettier and fit on page
  cat.just=list(c(0.6,1) , c(0,-5) , c(0,0) , c(1,1) , c(1,-4))
);

#Save the  Venn diagram
grid.newpage()
png("../Figures/venn_controlvheat_5sites_P0.05_04_26_2021.png")
grid.draw(venn.plot)
dev.off()

```
```{r venn_barshis_genes}
barshisgenes <- read.csv("../barshis2012_DElist.csv", skip = 4)
head(barshisgenes)
barshisgenes$ContigID <- barshisgenes$Ourcontigname
nrow(barshisgenes)
nrow(n12345)

shared <- subset(barshisgenes, ContigID %in% n12345$ContigID)

venn.plot <- draw.pairwise.venn(area1 = nrow(barshisgenes), area2 = nrow(n12345), cross.area = nrow(shared), category = c("Barshis genes", "Our genes"), fill = c("salmon", "skyblue"), lty = 1, cat.cex = 1.2, cat.pos = 1, cex = 3)


#Save the  Venn diagram
grid.newpage()
png("../Figures/venn_controlvheat_5sites_P0.05_barshisshared_allheatvcontrol_middleofvenn.png")
grid.draw(venn.plot)
dev.off()
```




